PoVaTiM: A Possibilistic Valid-Time Model \\ \vspace*{0.04truein}

### 2012-08-01 ijcis-pos-model/sources/timeRepresentation.tex
preprint version

This section is devoted to specify the representation of time within the framework of the possibility theory. First of all, the specification for a single ill-known temporal point will be explained. Then, the formal specification and the related constraints are given for an ill-known valid-time interval.

\subsection{\label{subsec:ill-known-point-rep}Ill-known time point}
An ill-known time point $X$ is a precise time point that, for some reason, is not fully specified. Note that $X$ has only one possible value but that value is unspecified.

@@

\begin{definition}
\label{def:ill-known-domain}
\textbf{Domain for an ill-known time point.}\\
Consider $\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants \emph{UNKNOWN} $=\left\lbrace 1/t, \forall t \in \T \right\rbrace$, \emph{UNDEFINED} $=\left\lbrace 0/t,\forall t \in \T \right\rbrace$ and \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$. The domain for an ill-known time point $X$ is given by: 
\begin{align}
\label{eq:ill-known-domain}
\Dom (X) =  \lbrace \Pow(\T) & \cup \mbox{\emph{UNKNOWN}} \\
\nonumber
&\cup \mbox{\emph{UNDEFINED}} \\
\nonumber
&\cup \mbox{\emph{NULL}} \rbrace
%
\end{align}
\end{definition}

\subsubsection{\label{subsubsec:ill-known-time-datatypes}Datatypes}
The data type for the representation of an ill-known time point allows the representation of the values shown in Table \ref{tbl:time-point-types}
%The datatypes for an  $X$ are shown in Table \ref{}.
\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:time-point-types}Values for the time point data type.}
\centerline{\small DIFFERENT VALUES FOR A TIME POINT}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{0.4\columnwidth} p{0.3\columnwidth}}\\
Subtype & Value & Representation \\
\hline
1 & A single time point &  $1/x, x \in \T$\\
2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t,\forall t \in \T \rbrace$\\
4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t,\forall t \in \T \rbrace$\\
5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
\hline\\
\end{tabular}
}
\end{samepage}

%Example of different uses of the ill-known time value.
\begin{example}
Consider a historical database with data from medieval diplomatic documents. The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the estimated time when the document was sent (field \emph{Date}). Table \ref{tbl:sample-time-point} contains some example data from this database. 
\end{example}

\vglue13pt
%\begin{table}[htbp]
\tcap{Sample of the historical database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c}\\
\textbf{ID} & Date\\
\hline
23454 & Unknown \\ 
34563 & 11/12/1204 \\
12211 & $\left[7/2/1204,30,30\right]$ \\
23455 & $\left[10,10/6/1204,20/6/1204,15 \right]$ \\
\hline\\
\end{tabular}
\label{tbl:sample-time-point}
}

@@

An ill-known time interval denoted by $\left[X, Y\right]$ is a precise time interval whose boundaries are not precisely known. %The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$. The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.

@@

%\item
\begin{definition}
\textbf{Semi-open time interval}: Only one of the two ill-known endpoints for the time interval $\left[X, Y\right]$ is unknown. Example \ref{ex:open-time-interval} and Figure \ref{fig:example-open-time-interval} illustrates a left open time interval.
% Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$. The solution adopted for this problem is explained in the following paragraph. 
\end{definition}
%\end{enumerate}

@@

\begin{definition}
\label{def:open-func}
$Open(C)$\\
Consider an ill-known value $T$, a binary relationship $B_r \in \left \lbrace \leq, <, >, \geq\right \rbrace$ and the constraint $C\stackrel{\triangle}{=}\left(R,T\right)$.
The function $\Open(C) = \left(I_p(C), I_n(C) \right)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.

@@

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/open-time-interval.pdf}
\label{fig:example-open-time-interval}
}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
\vspace*{13pt}
\end{samepage}

>>>>

This section is devoted to specify the representation of time within the framework of the possibility theory.

First of all, the specification for a single ill-known temporal point will be explained. Then, the formal specification and the related constraints are given for an ill-known valid-time interval.

\subsection{\label{subsec:ill-known-point-rep}Ill-known time point}
An ill-known time point $X$ is an atomic time point that, for some reason, is not fully specified.

Note that $X$ has only one possible value but that value is unspecified.

@@

\begin{definition}
\label{def:ill-known-domain}
\textbf{Domain for an ill-known time point.}\\
Consider $\tilde\Pow(\T)$ the set of all the possibility distributions over $\T$, and the three fuzzy constants:

\begin{itemize}

\item \emph{UNKNOWN} $=\left\lbrace 1/t, \forall t \in \T \right\rbrace$, ç
\item \emph{UNDEFINED} $=\left\lbrace 0/t,\forall t \in \T \right\rbrace$ and \item \emph{NULL} $=\lbrace 1/$ \emph{UNKNOWN}, $1/$ \emph{UNDEFINED} $\rbrace$. 
\end{itemize}

The domain for an ill-known time point $X$ is given by: 
\begin{align}
\label{eq:ill-known-domain}
\Dom (X) =  \lbrace \tilde\Pow(\T) & \cup \mbox{\emph{UNKNOWN}} \\
\nonumber
&\cup \mbox{\emph{UNDEFINED}} \\
\nonumber
&\cup \mbox{\emph{NULL}} \rbrace
%
\end{align}
\end{definition}

\subsubsection{\label{subsubsec:ill-known-time-datatypes}Datatypes}
The data type for the representation of an ill-known time point allows the representation of the values shown in Table \ref{tbl:time-point-types}.
%The datatypes for an  $X$ are shown in Table \ref{}.
\begin{samepage}
\vglue13pt
%\begin{table}[htbp]
\tcap{\label{tbl:time-point-types}Values for the time point data type.}
\centerline{\small DIFFERENT VALUES FOR A TIME POINT}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c p{0.4\columnwidth} p{0.3\columnwidth}}\\
Subtype & Value & Representation \\
\hline
1 & A single time point &  $1/x, x \in \T$\\
2 & A possibility distribution in the numeric domain & A fuzzy number or a fuzzy interval.\\
3 & An unknown value & \textbf{UNKNOWN}$= \lbrace1/t,\forall t \in \T \rbrace$\\
4 & An undefined value & \textbf{UNDEFINED}$= \lbrace 0/t,\forall t \in \T \rbrace$\\
5 & A null value & \textbf{NULL} $=\lbrace 1/$Unknown, 1/Undefined $\rbrace$\\
\hline\\
\end{tabular}
}
\end{samepage}

%Example of different uses of the ill-known time value.
\begin{example}
Consider a historical database with data from medieval diplomatic documents.

The following fields are stored: the digital identifier \emph{ID} which is the primary key  and the estimated time when the document was sent (field \emph{Date}).

Table \ref{tbl:sample-time-point} contains some example data from this database. 
\end{example}
\newpage
%\vglue13pt
%\begin{table}[htbp]
\tcap{Sample of the historical database}
%\centerline{\small DATA TYPES}
\vglue-6pt
\centerline{\small\baselineskip=13pt
\begin{tabular}{c c}\\
\textbf{ID} & Date\\
\hline
23454 & Unknown \\ 
34563 & 11/12/1204 \\
12211 & $\left[7/2/1204,30,30\right]$ \\
23455 & $\left[10,10/6/1204,20/6/1204,15 \right]$ \\
\hline\\
\end{tabular}
\label{tbl:sample-time-point}
}

@@

An ill-known time interval denoted by $\left[X, Y\right]$ is a time interval whose boundaries are not precisely known. %The ill-known time interval is defined by two ill-known  time points, namely $X$,$Y$. The evaluation of the ill-known time interval is given by equations \eqref{ill-known-pos},\eqref{ill-known-nec}.

@@

%\item
\begin{definition}
\textbf{Semi-open time interval}: Only one of the two ill-known boundaries for the time interval $\left[X, Y\right]$ is unknown. Example \ref{ex:open-time-interval} and Figure \ref{fig:example-open-time-interval} illustrates a left open time interval.
% Note that due to the ill-known constraints $C_{1}, C_{2}$ it is not possible a representation like $[$ \emph{UNKNOWN} $, Y]$ or $[X, $ \emph{UNKNOWN} $]$. The solution adopted for this problem is explained in the following paragraph. 
\end{definition}
%\end{enumerate}

@@

\begin{definition}
\label{def:open-func}
$Open(C)$\\
Consider an ill-known value $T$, a binary relationship $B_r \in \left \lbrace \leq, <, >, \geq\right \rbrace$ and the constraint $C\stackrel{\triangle}{=}\left(B_r,T\right)$.
The function $\Open(C) = \left(I_p(C), I_n(C) \right)$ provides both possibility and necessity measures for all the points in the open part of a semi-open ill-known time interval.

@@

\begin{samepage}
\vspace*{13pt}
\begin{center}
{
\includegraphics[scale=1]{./graphs/open-time-interval.pdf}

}
\end{center}
%\centerline{ \psfig{file=./graphs/Y-time-point.eps}}
\vspace*{10pt}
\fcaption{\label{fig:example-open-time-interval}Possibility distribution for $Y$, and possibility and necessity measures for the open ill-known point, $X$}
\vspace*{13pt}
\end{samepage}
